---
title: 'Course Project Part 1: Investigation of the Exponential Distribution'
author: 'Tatiana Lezhneva'
output:
  pdf_document: 
    fig_height: 3.5
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
cache = TRUE 
```

## Overview
 In this project we are going to investigate the Exponential Distribution (X ~ Exp(lambda)) and compare it with the Central Limit Theorem (CLT). The mean of the exponential distribution and standard deviation are 1/lambda. 
 
Simulation parameters: 

 * rate, lambda = 0.2 (for all of the simulations),
 
 * number of observation, n = 40,
 
 * number of simulations, sims = 1000.
 
## Simulations
```{r Simulations}
# set parameters for simulation
lambda <- 0.2
n <- 40
sims <- 1000
# run simulations
exp <- replicate(sims, rexp(n = n, rate = lambda))
# calculate mean of every simulation 
exp_means <- apply(x <- exp, MARGIN = 2, FUN = mean)
# plot histogram of means
hist(exp_means, breaks = 30, xlim = c(2,8), main = "Distribution of Means", xlab = "Mean")
```

## Results

### 1. Sample Mean versus Theoretical Mean

Lets calculate the mean of our simulations means distribution and plot it and on histogram. 
```{r 1, echo=FALSE}
hist(exp_means, breaks = 30, xlim = c(2,8), main = "Distribution of Means", xlab = "Mean")
abline(v = mean(exp_means), col = "red")
abline(v = 1/lambda, col = "green")
```

The sample mean `r (mean(exp_means))` (red line) is very close to the theoretical mean `r (1 / lambda)` (green line)

### 2. Sample Variance versus Theoretical Variance

```{r 2, echo=TRUE}
exp_var <- var(exp_means)
exp_sd <- sd(exp_means)
t_var <- (1/lambda)^2 / n
t_sd <- 1/(lambda * sqrt(n))
```
Variances and Standart Deviations are very close too:

standart deviation sample/ theoretical: `r (exp_sd)`/ `r (t_sd)`
 
variance sample/ theoretical:  `r (exp_var)` / `r (t_var)`

### 3. Distribution
The distribition of means looks like normal destribution curve:
```{r 3, echo=TRUE}
hist(exp_means, probability = TRUE, breaks = 30, xlim = c(2,8), 
     main = "Distribution of Means", xlab = "Mean") 
lines(density(exp_means), col = "green")
```

Lets looke at quantile-quantile plot:
```{r 4}
library(car)
qqPlot(exp_means, ylab = "destribution of means quantiles", main = "Normal Q-Q Plot")

```
Most of the points lie very close to a stright line, whith shows normal quantiles.

And finally we can count the 95% confidence interval of mean:
```{r}
mean(exp_means)+ c(-1,1) * qnorm(0.975) * sd(exp_means)/sqrt(n)
```



